---
title: "HW7"
author: "Bertrand Black"
date: "3/24/2021"
output: html_document
---
#### Open libraries 
```{r, echo=TRUE}
library(ggplot2)
library(MASS) 
```
### Simulation of my data
#### Add my data 
```{r, echo=TRUE}
z <- read.table("Athyrium_VT_Pinnae.csv",header=TRUE,sep=",", stringsAsFactors=FALSE)
names(z) <- list("myVar")
str(z)
summary(z)
```
#### Plot histogram of my data 
```{r, echo = TRUE}
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)
```

#### Plot unform, normal, exponetial, and gamma probability densities to determine which best fits the data set 
```{r, echo=TRUE}
# Plot `normal` probability density 
normPars <- fitdistr(z$myVar,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"]

p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)

meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$myVar),len=length(z$myVar))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
p1 + stat

# Plot `exponential` probability density 
expoPars <- fitdistr(z$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
p1 + stat + stat2

# Plot `uniform` probability density 
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
p1 + stat + stat2 + stat3

# Plot `gamma` probability density 
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
p1 + stat + stat2 + stat3 + stat4
```

### Test the beta probablity density on my data
```{r, echo=TRUE}
pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

#### Normal probability density best fits my data 
```{r, echo=TRUE}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
p1 + stat
```

#### New dataset simulation with same parameters as my data
```
normPars <- fitdistr(z2$myVar,"normal")
print(normPars)
     
#      mean          sd    
  10.7560000    2.8832731 
 ( 0.4077564) ( 0.2883273)
```
```{r, echo=TRUE}
z2 <- rnorm(n=50,mean=10.75)
z2 <- data.frame(1:50,z2)
names(z2) <- list("ID","myVar2")
z2 <- z2[z2$myVar2>0,]
print(z2)
str(z2)
summary(z2$myVar2)
```
#### Re-run the simulated data for comparison with the normal probablity density curve
```{r, echo=TRUE}
# Plot histogram of data 
p2 <- ggplot(data=z2, aes(x=myVar2, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p2)

# Add empirical density curve 
p2 <-  p2 +  geom_density(linetype="dotted",size=0.75)
print(p2)

# Get maximum likelihood parameters for `normal`
normPars2 <- fitdistr(z2$myVar2,"normal")
print(normPars2)
str(normPars2)
normPars2$meanML2 <- normPars2$estimate["mean"]
sdML2 <- normPars2$estimate["sd"]

meanML2 <- normPars2$estimate["mean"]
sdML2 <- normPars2$estimate["sd"]

xval2 <- seq(min(z2$myVar2),max(z2$myVar2),len=length(z2$myVar2))

stat2 <- stat_function(aes(x = xval2, y = ..y..), fun = dnorm, colour="red", n = length(z2$myVar2), args = list(mean = meanML2, sd = sdML2))
p2 + stat2 

```

#### My original data 
```{r, echo=FALSE}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
p1 + stat
```

#### How do the two histogram profiles compare? Do you think the model is doing a good job of simulating realistic data that match your original measurements? Why or why not?

* The simulated data does not fit as well with a normal distribution as my dataset. I chose to do fern pinnae lengths for my dataset. A continuous morphological dataset such as this is expected to form a normal distribution around the mean (10.75). The simulated dataset preformed well but was not as close of a fit. 

* I do not think the simulated data preformed as well as the because with such a small sample size the randomly generated numbers by chance were not as well distributed as my original dataset. With more a greater sample size it would probably improve.


